---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '#'
---

# Arguments {#sec-args}

```{bash}
#| label: setup
#| include: false
rm -Rf tmp
whoami
mkdir tmp
echo "This is my file" > my_file.txt
```

```{r}
#| label: common
#| include: false
# Sys.setenv('QUARTO_PROJECT_DIR' = 'data')
# Sys.getenv('QUARTO_PROJECT_DIR')
source("_common.R")
```

```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

Arguments in Unix/Linux commands are values or data passed to commands for processing. Unlike options, which modify the behavior of commands, arguments typically specify what the command should *act upon*. This typically includes filenames, user names, data values, or other kinds of information the command needs to execute its task.
 
## [Anatomy]{style="font-size: 1.05em; font-weight: bold;"}

A Unix command can be broken down into the command name, followed by its options (which we'll address in the next chapter), and then its arguments:

```{bash}
#| eval: false
command argument1 argument2 ... argument
```

## [Argument Types]{style="font-size: 1.05em; font-weight: bold;"} 

Below are a variety of command arguments types. This is not an exhaustive list, but includes many of the commands and arguments you'll encounter on a regular basis.

### Direct arguments

**Direct arguments** are the most straightforward type of arguments. They are typically the names of files or directories on which commands operate. 

**Example** 

In the command `cat my_file.txt`, `my_file.txt` is a direct argument to the <code>@sec-cat</code> command, telling it which file to display on the standard output.

```{bash}
cat my_file.txt
```

###  Indirect arguments

**Indirect arguments** are arguments that might specify additional information that commands need to complete their tasks.

**Example** 

The `file` search pattern for the <code>@sec-grep</code> command is an example of an indirect command, and `myfile.txt` is the direct argument.

```{bash}
grep file my_file.txt
```

Most commonly, arguments are the names of **files and directory names** on which the command will operate.

**Example**

`my_file.txt` and `tmp/tmp_file.txt` are arguments representing the source and destination locations to move with <code>@sec-mv</code>, respectively.

```{bash}
#| eval: true
mv my_file.txt tmp/tmp_file.txt
```

Commands related to user management might take **user and group names** names as arguments.

**Example**

<code>@sec-chown</code> changes the ownership of `file` to `user` and `group`.

```{bash}
#| eval: false
chown user:group myfile.txt
```

### Command Targets

Some commands take **other commands as arguments**.

**Example**

For example, <code>@sec-sudo</code> command runs command with superuser privileges.

```{bash}
#| eval: false
sudo vi path/to/file.config
```

### Data Values

Commands might take **data values** as arguments for processing. 

**Example**

In `echo Hello, World!`, `Hello, World!` is the argument value that <code>@sec-echo</code> prints to the terminal.

```{bash}
#| eval: true
echo Hello, World!
```

### Order and Position

For many commands, the **order** of the arguments is significant. 

**Example** 

`tmp/tmp_file.txt` is the first argument (indicating the file to copy from), and `new_file.txt` is the second argument (indicating where to copy the file to). 

```{bash}
#| eval: true
cp tmp/tmp_file.txt new_file.txt
```

Reversing these arguments would result in a completely different operation.

Arguments that contain **spaces** must be quoted or escaped, so the shell understands them as a single argument rather than multiple arguments. 

**Example** 

To copy the contents of the new `new_file.txt` to `'my new file.txt'`, you would use:

```{bash}
#| eval: true
cp new_file.txt 'my new file.txt'
```

### Command Substitution

The **output of a command can be used as an argument for another command** using backticks (\` \`) or `$( )`. 

**Example** 

`echo $(grep file 'my file 2.txt')` uses the output of the <code>@sec-grep</code> command as an argument for <code>@sec-echo</code>:

```{bash}
#| eval: true
echo $(grep file 'my new file.txt')
```

### Variables as Arguments

**Environment variables** can be used as arguments in commands. 

**Example** 

`echo $HOME` prints the path to the user's home directory, where `$HOME` is an argument that the <code>@sec-echo</code> command interprets:

```{bash}
#| eval: false
echo $HOME
# /Users/username
```

## [Recap]{style="font-size: 1.05em; font-weight: bold;"} 


Understanding the nuances of Unix arguments is crucial for crafting precise and effective commands, allowing users to leverage the full power of the Unix command line for a wide array of tasks.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```


```{bash}
#| label: cleanup
#| include: false
rm -Rf tmp
rm new_file.txt
rm 'my new file.txt'
```