---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '# '
---

# Arguments {#sec-args}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
# Sys.setenv('QUARTO_PROJECT_DIR' = 'data')
# Sys.getenv('QUARTO_PROJECT_DIR')
source("_common.R")
```


```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

Arguments follow the command and provide additional information or specify targets for the command, such as file names and paths. 

```bash
$ command [argument]
```

We'll use the <code>@sec-ls</code> command to list the directory contents of the root directory (accessed with `/`):

```{bash}
ls /
```

In the example above, `/` is the argument passed to the `ls` command.

## Quotes

For example, in <code>@sec-cat</code> `myfile.txt`, the `myfile.txt` file is an argument to the `cat` command, specifying which file it should display.

```{bash}
cat myfile.txt
```

```{bash}
cat 'myfile.txt'
```


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
