---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '# '
---

# Basics  {#sec-basic-cmnds .unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

The next three section introduce common commands that will help users navigate folders (or directories), manipulate and manage files, check system processes, and use or create variables. These sections focus on executing single commands at a Terminal. In the next section we'll dive into the Unix/Linux syntax (i.e., arguments, options, and pipes).

#### Basic Unix/Linux Commands

[Directories](dirs.qmd) covers how to navigating the file system. 

[Files](dirs.qmd) discusses common file operations.

[System](sys.qmd) This chapter covers common system monitoring commands in Unix/Linux: `top` and `htop` for real-time process monitoring, `free` for memory usage, `df` for disk space usage, and `du` for directory and file size. These tools help efficiently manage and troubleshoot system performance and resource usage.

[Variables](vars.qmd) dives into the definition, types, and use of variables. Environment variables (`PATH`, `HOME`, `USER`), commands used with variables (`printenv`, `set`, `unset`, `export`, `env`), scripts, integer and strings, and best practices.

<!--

#### Advanced Topics

- **Scripting and Automation**: Basics of writing shell scripts to automate routine tasks.

- **Customizing the Environment**: How to customize the shell experience, including themes, prompts, and aliases.

- **Security Practices**: Basic security practices using terminal commands, such as setting file permissions and managing user accounts.

#### Practical Exercises

- **Hands-On Examples**: Provide step-by-step exercises to practice basic and intermediate commands.

- **Project Ideas**: Offer project suggestions that involve using terminal commands to solve real-world problems, like data backup scripts or a basic server setup.

-->

## Recap 

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```