---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '# '
---

# Files and Folders  {.unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

#### `mkdir`

**`mkdir` (Make Directory)** builds a new folder wherever you tell it to, like making a new folder for projects in `/bin` (although you typically wouldn't do that there)

#### `cp`

**`cp` (Copy)** duplicates files or folders. For instance, copying a tool from `/bin` to your desktop, but usually, you leave `/bin` tools where they are.

#### `mv`

- **`mv` (Move)**: Taking a book from your desk and putting it on the shelf, `mv` moves files or folders from one place to another. It's like moving a misplaced tool back into `/bin`.

#### `rm`

**`rm` (Remove)**: Throwing away an old, tattered magazine, `rm` deletes files or folders. Be cautious, as this is like shredding documents—no easy way to get them back.

#### `ln`

**`ln` (Link)**: Creating a shortcut on your desktop to a frequently used program, `ln` creates links to files, making them accessible from multiple places without duplicating them.
