---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Glossary {.unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

## Bash {#sec-bash}

[Bash](https://www.gnu.org/software/bash/),  short for Bourne Again SHell, is a command line interface and scripting language for operating systems, enabling direct command input and task automation. Originally created for the GNU project and known for its flexibility and powerful features, Bash is the standard shell on many Linux distributions and was the default shell in the Terminal on macOS until the Catalina release.

## Quarto {#sec-quarto}

[Quarto](https://quarto.org/) is an open-source scientific and technical publishing framework designed to work with R, Python, Julia, Observable JavaScript, and more, making it a versatile tool for data scientists, researchers, and anyone involved in data analysis.

## YAML {#sec-yaml}

YAML: [YAML](https://yaml.org/) is a human-friendly data format for configuration files and data exchange, using key-value pairs, lists, and indentation to organize data.

```yaml
key: value
  key: value
```

It's readable and easily parsed by machines, making it popular for application configuration and data sharing.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```

