---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
editor: 
  markdown: 
    wrap: 72
---

# Welcome! {.unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This book is under development. Thank you for your patience.
  "
)
```

Welcome to the beginning of your Unix and Linux journey! This book aims
to provide a clear understanding of the concepts and commands of
Unix/Linux, even for non-technical audiences. Whether you are a student,
a professional from a non-IT field, or someone curious about operating
systems, this book will guide you through the essentials of Unix/Linux.

<!-- ## What This Book Is Not -->

<!-- ## Audience and Prerequisites -->

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
