---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Introduction {.unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

This is a book created from markdown and executable code.

> "*Let us change our traditional attitude to the construction of programs: Instead of imagining that our main task is to instruct a computer what to do, let us concentrate rather on explaining to human beings what we want a computer to do.*" - [Donald Knuth. "Literate Programming (1984)" in Literate Programming. CSLI, 1992, pg. 99.](http://www.literateprogramming.com/index.html)

```{bash}
whoami
```

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
