---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Preface {#sec-preface .unnumbered}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

Unix and Linux are powerful, versatile, and widely used operating
systems that run on millions of devices worldwide, from smartphones to
supercomputers. Learning about Unix/Linux can open doors to many career
opportunities and help you better understand the digital world around
you.

This book is structured to gradually introduce you to the Unix/Linux
environment, starting from the basics to more complex concepts. Here`s
what each section covers:

**Introduction**

The introduction acquaints you with some background that formed the
landscape of Unix/Linux systems. It explains the Bash shell, a
command-line interface (CLI), which is the gateway to leveraging the
full potential of Unix/Linux systems.

**Set-Ups**

Before diving into the commands and scripts, setting up your
Unix/Linux environment is crucial. This section guides you through
various options for setting up Unix/Linux on your machine, be it
natively or through a virtual machine.

**Syntax**

Understanding the syntax is vital for effectively
communicating with the Unix/Linux system. This section demystifies the
structure of commands, including how to differentiate between commands,
arguments, and options and manage inputs and outputs. This knowledge is
critical to executing tasks efficiently in the command line. 

@sec-pipes are a cornerstone of Unix/Linux productivity, enabling
the output of one command to serve as the input to another. This section
covers how to combine commands using `stdout` and `stdin`, allowing for
robust command chains that can perform complex tasks with a simple
syntax. Understanding pipes unlocks a higher level of command-line
efficiency and is a step towards advanced Unix/Linux usage.

**Files and Directories**

Navigating and managing files and folders are
daily tasks for Unix/Linux users. @sec-dirs covers essential commands
such as `cd` for changing directories, `pwd` to print the current
directory, `ls` for listing files, and `mkdir` for making directories.

@sec-files chapter covers file manipulation commands like `cp` (copy), `mv`
(move), `rm` (remove), and `less` for viewing file content, ensuring you
can organize and manage your file system effectively. These commands are
the building blocks for more complex operations in Unix/Linux.

**Manipulating Text**

**Tex Commands**: Unix/Linux systems are renowned for their powerful
text manipulation capabilities. This section introduces commands such as
`cat` for displaying file contents, `grep` for searching within
files, `sort` for sorting data, `uniq` for filtering unique lines, and
`cut`, `paste`, and `join` for editing files. Mastering these commands
will allow you to handle and process text data efficiently.

**Text Editors:** This section covers nano, vim, and emacs.

**Shell Scripts (in development)**

Shell scripting is a powerful tool for automating repetitive tasks in
Unix/Linux. This section introduces you to writing your shell scripts,
covering the basics of script creation, execution, and debugging. You`ll
learn how to automate simple tasks, making your Unix/Linux experience
more productive and enjoyable.

**Format**: Shell scripts can be written in various formats, depending
on your shell (e.g., Bash, Zsh). This section delves into the
differences between these formats, guiding you on writing compatible
scripts that can run across different Unix/Linux systems.

**Permissions**: Understanding file permissions is crucial to managing
the security of your files and directories in Unix/Linux. This section
explains how Unix/Linux permissions work and teaches you how to set and
modify permissions to protect your data and system from unauthorized
access.

By the end of this book, you'll have a solid understanding of the
Unix/Linux operating system. You'll be equipped to navigate, manage
files, write scripts, and set permissions confidently. Whether you're
looking to enhance your career prospects, manage your Unix/Linux
systems, or simply satisfy your curiosity, this book will be your
companion on a fascinating journey into the world of Unix/Linux.

## <!--# What This Book Is Not   -->

## <!--# Audience and Prerequisites -->

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```