---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Preface {#sec-preface .unnumbered}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

Unix and Linux are powerful, versatile, and widely used operating systems that run on millions of devices worldwide, from smartphones to supercomputers. Learning about Unix/Linux can open doors to many career opportunities and help you better understand the digital world around you.

This book is structured to gradually introduce you to the Unix/Linux environment, starting from the basics to more complex concepts, then covering multiple case studies.


## [Book outline]{style="font-size: 1.05em; font-weight: bold;"}

The first four sections of this manual cover concepts and materials for users who are new to Unix/Linux. If you're familiar with Unix/Linux command-line tools and regular expressions, feel free to skip these chapters.

### Introduction

The <strong>@sec-intro</strong> acquaints you with some background that formed the landscape of Unix/Linux systems. It explains the Bash shell, a command-line interface (CLI), which is the gateway to leveraging the full potential of Unix/Linux systems.

### Set-Ups

Before diving into the commands and scripts, setting up your Unix/Linux environment is crucial. <strong>[Set-ups](setups.qmd)</strong> guides you through various options for setting up <strong>@sec-sh-term</strong>, <strong>@sec-vms</strong>, and <strong>@sec-quarto</strong> documents.

### The Basics

Navigating and managing files and folders are daily tasks for Unix/Linux users. <strong>@sec-dirs</strong> covers essential commands such as `cd` for changing directories, `pwd` to print the current directory, `ls` for listing files, and `mkdir` for making directories.

<strong>@sec-files</strong> chapter covers file manipulation commands like `cp` (copy), `mv` (move), `rm` (remove), and `less` for viewing file content, ensuring you can organize and manage your file system effectively. These commands are the building blocks for more complex operations in Unix/Linux.

The <strong>@sec-sys</strong> chapter goes over common system commands like `top` for monitoring processes in real-time, `ps` for listing currently running processes, `kill` to stop a process, `df` (Disk Free) to display disk space usage on all mounted filesystems, and `du` (Disk Usage) to estimate file space usage. These commands help you manage the system's processes and resources efficiently, ensuring you have a clear view of resource allocation and consumption.

### Syntax

Understanding the syntax is vital for effectively communicating with the Unix/Linux system. The <strong>[Syntax](syntax.qmd)</strong> section demystifies the structure of commands, including how to differentiate between <strong>@sec-commands</strong>, <strong>@sec-args</strong>, and <strong>@sec-options</strong> and manage inputs and outputs. This knowledge is critical to executing tasks efficiently in the command line.

<strong>@sec-pipes</strong> are a cornerstone of Unix/Linux productivity, enabling the output of one command to serve as the input to another. <strong>[Syntax](syntax.qmd)</strong> also covers how to combine commands using `stdout` and `stdin`, allowing for robust command chains that can perform complex tasks with a simple syntax. Understanding pipes unlocks a higher level of command-line efficiency and is a step towards advanced Unix/Linux usage.

### Manipulating Text (in development)

The <strong>[Text](text.qmd)</strong> section covers using <strong>@sec-symbols-patterns</strong> to build regular expressions (regex). Unix/Linux systems are renowned for their powerful text manipulation capabilities. The <strong>@sec-txt-cmds</strong> chapter introduces commands such as `cat` for displaying file contents, `grep` for searching within files, `sort` for sorting data, `uniq` for filtering unique lines, and `cut`, `paste`, and `join` for editing files. Mastering these commands will allow you to handle and process text data efficiently. <strong>[Text Editors](text_editors.qmd)</strong> covers nano, vim, and emacs.

### Shell Scripts (in development)

Shell scripting is a powerful tool for automating repetitive tasks in Unix/Linux. This section introduces you to writing your shell scripts, covering the basics of script creation, execution, and debugging. You'll learn how to automate simple tasks, making your Unix/Linux experience more productive and enjoyable.

The <strong>[Format](format.qmd)</strong> chapter delves into the differences between these formats, guiding you on writing compatible scripts that can run across different Unix/Linux systems. <strong>[Permissions](permissions.qmd)</strong> explains how Unix/Linux permissions work and teaches you how to set and modify permissions to protect your data and system from unauthorized access.

### Use cases (in development)

## [Data files]{style="font-size: 1.05em; font-weight: bold;"}

The data files used in this book are documented in [`data/README.md`](https://github.com/mjfrigaard/fm-unix/tree/main/data/README.md).

## [Recap]{style="font-size: 1.05em; font-weight: bold;"}

By the end of this book, you'll have a solid understanding of the Unix/Linux operating system. You'll be equipped to navigate, manage files, write scripts, and set permissions confidently. Whether you're looking to enhance your career prospects, manage your Unix/Linux systems, or simply satisfy your curiosity, this book will be your companion on a fascinating journey into the world of Unix/Linux.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
