---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
---

# Set-Ups {#sec-setups .unnumbered}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
# Sys.setenv('QUARTO_PROJECT_DIR' = 'data')
# Sys.getenv('QUARTO_PROJECT_DIR')
source("_common.R")
```


```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

This section details the practical aspects of preparing your environment to work with Linux systems. It is designed to guide you through various setup processes, catering to different preferences and requirements. We’ll cover working with shells and terminals, setting up a virtual machine, and using Quarto documents to execute Bash commands.

### Shells and Terminals 

<strong>@sec-sh-term</strong> covers the differences between shells and terminals, and some common options for both.

### Virtual Machines 

<strong>@sec-vms</strong> covers how to run Linux on a virtual machine.

### Quarto 


<strong>@sec-quarto</strong> covers how to setup up the [Quarto](https://iterm2.com/) publishing system.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
