---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
---

# Set-Ups {#sec-setups .unnumbered}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
# Sys.setenv('QUARTO_PROJECT_DIR' = 'data')
# Sys.getenv('QUARTO_PROJECT_DIR')
source("_common.R")
```


```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

This book section details the practical aspects of preparing your environment to work with Unix/Linux systems. It is designed to guide readers through various setup processes, catering to different preferences and requirements. We’ll cover using Quarto documents to execute Bash commands, working directly on a macOS terminal, and setting up a virtual machine.

<strong>@sec-terminals</strong> covers how to setup up [iterm2](https://iterm2.com/).

<strong>@sec-quarto</strong> covers how to setup up the [Quarto](https://iterm2.com/) publishing system.

<strong>@sec-vms</strong> covers how to run Unix on a virtual machine.

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```