---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    
---

# Set-Ups {#sec-setups .unnumbered}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
# Sys.setenv('QUARTO_PROJECT_DIR' = 'data')
# Sys.getenv('QUARTO_PROJECT_DIR')
source("_common.R")
```


```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

This section details the practical aspects of preparing your environment to work with a Linux system. We’ll cover setting up a virtual machine and working with shells and terminals. It is designed to guide you through various setup processes, catering to different preferences and requirements.

### Virtual Machines 

<strong>@sec-vms</strong> covers how to run Linux on a virtual machine.

### Shells 

<strong>@sec-shells</strong>

### Terminals 

<strong>@sec-terminals</strong>

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
