---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Set-Up {.unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

## Using Bash in Quarto 

Follow the instructions on the [Quarto website](https://quarto.org/docs/computations/execution-options.html#shell-commands) to configure your HTML documents to run Bash code chunks. 

### Document header

Specify the `engine` field of in the YAML header of the `.qmd` file:

```yaml
---
title: "Using Bash"
engine: knitr
---
```

You can also add any `knitr` chunk options: 

```yaml
knitr:
  opts_chunk: 
    collapse: true
```

### Code Chunks

The code chunks function like R or Python code chunks: 

```{{bash}}
echo "foo" 
```

You caan also specify the code chunk options with the hash-pipe:

```{{bash}}
#| code-fold: show
#| code-summary: 'show/hide echo'
echo "foo" 
```

```{bash}
#| code-fold: show 
#| code-summary: 'show/hide echo'
echo "foo"
```

## Terminal (macOS)

## Virtual Machines

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```