---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Set-Up {.unnumbered}

```{r}
#| label: setup
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

##  Quarto 

[Quarto](https://quarto.org/) is an '*open-source scientific and technical publishing system*' that enables users to create dynamic and beautiful documents, reports, presentations, and dashboards. 

```{r}
#| label: co_box_knuth
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "b",
  look = "default", 
  hsize = "1.10", 
  size = "1.0", 
  header = "Literate Programming", 
  fold = TRUE,
  contents = "
> '*Let us change our traditional attitude to the construction of programs: Instead of imagining that our main task is to instruct a computer what to do, let us concentrate rather on explaining to human beings what we want a computer to do.*' - [Donald Knuth. Literate Programming (1984) in Literate Programming. CSLI, 1992, pg. 99.](http://www.literateprogramming.com/index.html)")
```


@sec-quarto documents are built to work with various programming languages, including Unix commands. 

### YAML header

Quarto documents are written in markdown, a lightweight markup language that’s easy to write and read. This simplicity allows authors to focus on their content rather than formatting. Quarto documents contain a @sec-yaml header for configuring various options. To run Bash commands, specify `knitr` in the `engine` field of in the YAML header of the `.qmd` file, and any additional key-value pairs:

```yaml
---
title: "Using Bash"
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---
```

### Bash Code Chunks

Introduced in 1989, Bash has become the default command-line interface or "shell" for most Linux distributions.[^bash-macos] We'll use Bash commands and scripts to preproceess data files, render Quarto documents, and automate other operations in the data analysis pipeline, from downloading and ingesting data to creating and deploying analysis outputs.  

[^bash-macos]: Bash was the default command-line interface for Apple's macOS (which is Unix-based) until the transition to `zsh` as the default shell in macOS Catalina.

One of the powerful aspects of Quarto is its ability to integrate executable code chunks into Markdown documents (in fact, this entire book was created using markdown and executable code chunks).

```{{bash}}
echo "foo" 
```

You can also specify the code chunk options with the hash-pipe (`#|`):[^quarto-bash] 

[^quarto-bash]: Follow these instructions on the [Quarto website](https://quarto.org/docs/computations/execution-options.html#shell-commands) to configure your HTML documents to run Bash code chunks.

```{bash}
#| code-fold: show 
#| code-summary: 'show/hide echo'
echo "foo"
```

```{{bash}}
#| code-fold: show
#| code-summary: 'show/hide echo'
echo "foo" 
```


## Terminal (macOS)

## Virtual Machines

```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```