---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: ' '
---

# Syntax {.unnumbered}

```{bash}
#| label: setup
#| include: false
touch myfile.txt
echo 'This is my file' > myfile.txt
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_rev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "o",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is being revised. Thank you for your patience."
)
```

In the previous section we covered some of the basic common commands you'll use in a Linux-like operating systems. This section dives deeper into the syntax--the commands, arguments, and options--you'll type into the terminal or write in a shell script. We'll also discuss creating variables and  how to chain commands together using pipes

### Commands 

### Arguments 

### Options 

### Pipes 

### Variables 


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```
