---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
---

# Text Commands {#sec-txt-cmds}

```{bash}
#| label: setup
#| include: false
whoami
```


```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

The Unix and Linux operating systems thrive on simplicity and efficiency, principles elegantly manifested in text handling. Central to this ecosystem are the commands explicitly designed for text manipulation—tools such as `cat`, `sort`, `cut`, and others form the core of text processing activities.

### The Text Stream

Unix/Linux conceptualizes text as a stream, a continuous sequence of characters that can be manipulated in real-time. Streams are crucial for understanding how Unix/Linux commands process text. A text stream can originate from files, input devices, or even the output of other commands. Treating text as a steady stream of inputs offers a versatile and powerful method for text manipulation.


<!--

- **`cat`**: Beyond simply displaying the content of files, `cat` is often used to concatenate several text files into one and to work with text streams.
- **`sort`**: As the name suggests, `sort` organizes text lines into a specified order, proving invaluable for data analysis and organization tasks.
- **`cut`**: This command extracts sections from each line of input, making it essential for parsing and processing delimited data.
- **`grep`**: A powerful search tool, `grep` filters text according to patterns, enabling complex data extraction and analysis operations.
- **`awk`**: An entire programming language in itself, `awk` is designed for pattern scanning and processing, offering sophisticated text manipulation capabilities.
- **`sed`**: The stream editor (`sed`) transforms text in a stream, allowing for powerful pattern matching and substitution.

Each of these commands leverages stdin and stdout, enabling them to be combined in pipelines that filter, transform, and manipulate text in complex ways with just a few keystrokes. This section aims to demystify these commands, providing a foundation for understanding and utilizing Unix's rich text manipulation capabilities.

By mastering these commands, you gain the ability to effortlessly manipulate text streams, transforming raw data into structured information. This knowledge is not just a technical skill but a gateway to the deeper philosophy of Unix, where simplicity and efficiency reign supreme.
-->

## Text Manipulation 

These commands embody the Unix philosophy of '*do one thing and do it well*' and demonstrate the system’s power in processing text streams. This section will explore these fundamental commands, illustrating how they easily interact with text streams, standard input (`stdin`), and standard output (`stdout`) to perform complex text manipulations.

```{r}
#| label: co_box_stdin_stdout
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "b",
  look = "default", 
  hsize = "1.10", 
  size = "1.05", 
  header = " Refresher: Standard Input and Standard Output", 
  fold = TRUE,
  contents = "Two key concepts in Unix text processing are standard input (`stdin`) and standard output (`stdout`). `stdin` is the default input stream, which often comes from the keyboard or the output of another command. `stdout` is the default output stream, typically the terminal screen. Many Unix commands read from `stdin` when no file is specified and write to `stdout`, allowing the output of one command to become the input of another. This design facilitates the chaining of commands (piping) to perform complex operations in a streamlined manner.

-   **Input** generally refers to the data fed into a command, which can come from `stdin` or be specified as arguments.

-   **Output** is the data produced by a command, displayed on `stdout` unless redirected.")
```

### `echo` 

The `echo` command prints its arguments to the standard output. It's commonly used in scripts and on the command line to display messages or variables.

### `cat` 

**`cat` (Concatenate)**: `cat` displays the content of files straight to your screen, useful for checking what's in a text file quickly.

This is similar to printing a file and laying out the pages on the floor to see them all at once.

### `grep` 

**`grep`** stands for "global regular expression print" and it reads from `stdin` or a list of files and outputs the lines that contain matches for a specified pattern.


### `sort` 

**`sort`**: `sort` arranges the lines in a text file into order:


Similar to organizing our stack of papers alphabetically.

### `uniq` 

**`uniq`**: `uniq` helps by removing duplicate lines from a file, making sure every line is unique.

After accidentally printing duplicates of a document, you remove the extra copies.

### `cut` 

**`cut`**: `cut` extracts specific parts of lines in a file, like cutting out columns of text.

If you only want the dates from a list of events, you might physically cut them out of the paper. 

### `paste` 

**`paste`**: `paste` combines lines from multiple files side by side:

`paste` is like taking snippets of text from different documents and sticking them together into one.

### `join` 

**`join`**: `join` merges lines from two files based on a common field.


If you have two lists with common information, you might merge them based on what matches. 

### `comm` 

**`comm` (Compare)**: `comm` compares two sorted files line by line to see what items appear on both, only on one, or the other.

### `diff` 

**`diff` (Difference)**: Highlighting what changes have been made between your draft and the final copy of a letter, `diff` shows the differences between two files.

### `wc` 

`wc` (**Word Count**): `wc` counts the number of lines, words, and characters in the given input. If a file name is provided, it performs the count on the file; otherwise, it reads from the standard input.

### `xargs` 

`xargs` builds and executes command lines from standard input (`stdin`). It's most often used in combination with other commands through pipes. `xargs` takes input from a pipe and passes it as arguments to another command, allowing for powerful command-line operations that process a list of inputs iteratively.

## Recap 

These analogies help demystify what can feel like complex commands, tying them back to everyday actions and decisions.


```{r}
#| label: git_contrib_box
#| echo: false
#| results: asis
#| eval: true
git_contrib_box()
```