---
engine: knitr
knitr:
  opts_chunk: 
    collapse: true
---

# Variables {#sec-vars}

```{bash}
#| label: setup
#| include: false
whoami
```

```{r}
#| label: common
#| include: false
source("_common.R")
```

```{r}
#| label: co_box_dev
#| echo: false
#| results: asis
#| eval: true
co_box(
  color = "r",
  look = "default", 
  hsize = "1.25", 
  size = "1.00", 
  header = "Caution", 
  fold = FALSE,
  contents = "This section is under development. Thank you for your patience.
  "
)
```

This chapter covers the essential aspects of variables in Unix/Linux environments.

## [Unix/Linux Variables]{style="font-size: 1.05em; font-weight: bold;"} 

Definition and significance of variables in Unix/Linux.

Overview of environment variables vs. shell variables.

## [Types of Variables]{style="font-size: 1.05em; font-weight: bold;"} 

Variables come in two flavors--Shell and Environment. 

### Shell Variables

Local to the shell instance.

### Environment Variables

Available system-wide and inherited by child processes.

Differences and use cases.

## [Setting Variables]{style="font-size: 1.05em; font-weight: bold;"} 

### Setting Shell Variables

Basic syntax for setting variables (`varname=value`).

### Setting Environment Variables

Basic syntax for exporting environment variables (`export VARNAME=value`).

## [Using Variables]{style="font-size: 1.05em; font-weight: bold;"} 

### Using Shell Variables

Accessing variable values (`echo $varname`).

Examples of common shell variables.

### Using Environment Variables

How to make environment variables persistent (`.bashrc`, `.profile`, etc.).

Examples of critical environment variables (`PATH`, `HOME`, `USER`).

## [Variables, Commands and Utilities]{style="font-size: 1.05em; font-weight: bold;"} 

- `printenv`: Displaying environment variables.

- `set`: Viewing shell and environment variables.

- `unset`: Deleting variables.

- `export`: Marking variables to be exported to child processes.

- Using `env` for running programs in a modified environment.

### Variables in Scripts

- How to use variables to customize scripts.

- Passing variables to scripts via command line.

- Reading and modifying variables within scripts.

- Example script demonstrating variable use.

### Advanced Variable Features

- Arrays and associative arrays in Bash.

- Integer operations and manipulating string variables.

- Special variables (`$0`, `$1`, `$?`, `$$`, `$@`).

### Security and Best Practices

- Security implications of improperly handling variables.

- Best practices for naming and using variables.

- Avoiding common pitfalls (e.g., overwriting critical environment variables).

## [Recp]{style="font-size: 1.05em; font-weight: bold;"} 

- Recap of key points covered in the chapter.

- Additional resources for deeper exploration (books, online tutorials, communities).
